### 2024-03-12 paper/paper.qmd
did data section, made tables, fixed some minor and major issues with formatting and references

---
title: "Homelessness and Drug Toxicity"
author: 
  - Gavin Crook
  - Samarth Rajani
  - Mohid Sharif
thanks: "Code and data are available at: https://github.com/MohidSharif/Drug-Toxicity-and-Homelessness.."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(here)
# Turn off echo for all code, disables knitting code chunks
knitr::opts_chunk$set(echo = TRUE)
# Turn off warning messaging in knitted paper file
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
# Import needed libraries
library(kableExtra)

cleaned_data <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

We obtained our data from the City of Toronto **opendatatoronto** database portal, using the 'opendatatoronto' package [@citeopendatatoronto] and the statistical programming language **R** [@citeR]. I used the **tidyverse** package for data manipulation [@citetidyverse] and **kableExtra** for table formatting [@citekableExtra]. The header includes two lines of code **"usepackage{float}"** which allows the use of float in R markdown and the line **"floatplacement{figure}{H}"** [@citeholdposition] which keeps the tables and figures locked in the specific place where they are written in R markdown.

This data set records reports of homeless deaths across Toronto and records the details of the deceased. The data records the **year of death, cause of death, age group, gender, and number of deaths** for every report. The data classifies cases by age group rather than specific ages, age groups are grouped by 20 year age gaps (E.g. "20 to 39 Years") except for the first group which is **"<20"** and the last group which is **"60+"**. Gender is identified and reported as **"Male"** or **"Female"**. Year of death is simply as the year the deaths were reported, starting from **2017** up until **2023**. Cause of deaths are classified as follows: **"Accident", "Cancer", "Cardiovascular Disease", "Covid-19", "Drug Toxicity", "Homicide", "Pneumonia", "Other", "Suicide", and "Unknown/Pending"**. And the number of deaths is simply the number of deaths provided in that report.

Since we are only interested in deaths due to drug toxicity, we simplified the data and created two datasets. First the data was cleaned to remove any **"Unknown"** or **empty** values. Then causes of death due to **Drug Toxicity** were isolated and all others were removed from the dataset. We then created two datasets, one highlighting the number deaths per year for each gender and one for the number of deaths per year for each age group. Since case in the data set reported multiple deaths, therefore all counts from each report had to be added to their respective grouping to create a new death count variable for the two datasets. Using these new datasets we can now compare and analyze the death trend oover the years for each gender and age group.

@tbl-sample1 shows the data associated with cases reported as "death due to drug toxicity".This data shows us year, age group, gender and count of each reported case of drug toxicity.

```{r}
#| label: tbl-sample1
#| tbl-cap: Age Group Deaths Due to Drug Toxicity
#| echo: false

# Import knitr
library(knitr)

# Read the data into a variable
age_group_data <- read_csv(here::here("data/analysis_data/age_group_deaths_data.csv"))
gender_data <- read_csv(here::here("data/analysis_data/gender_deaths_data.csv"))

# Filter the data to be specific to data relating to drug txocity causes of death
age_group_data |> 
  select(Year, Age_group, Deaths) |>
# Include 10 listings in the table
  slice(1:8) |> 
# Display the table with appropriate column names and format properties
  kable(
    col.names = c("Year", "Age Group", "Deaths"),
    digits = 1,
    booktabs = TRUE,
    linesep = "") |> 
    kable_styling(latex_options = "HOLD_position")

```

@tbl-sample2 shows the data associated with cases reported as "death due to drug toxicity".This data shows us year, age group, gender and count of each reported case of drug toxicity.

```{r}
#| label: tbl-sample2
#| tbl-cap: Gender Deaths Due to Drug Toxicity
#| echo: false

# Filter the data to be specific to data relating to drug txocity causes of death
gender_data |> 
  select(Year, Gender, Deaths) |>
# Include 10 listings in the table
  slice(1:8) |> 
# Display the table with appropriate column names and format properties
  kable(
    col.names = c("Year", "Gender", "Deaths"),
    digits = 1,
    booktabs = TRUE,
    linesep = "") |> 
    kable_styling(latex_options = "HOLD_position")

```

Talk more about it.

Talk way more about it.

```{r}
#| label: figure1
#| fig-cap: Deaths Per Year for Age Group
#| echo: false
#| warning: false
#| message: false

# Aggregate data to get the total number of deaths for each year and age group
total_deaths <- cleaned_data %>%
  group_by(Year_of_death, Age_group) %>%
  summarise(Total_Deaths = sum(Count),.groups = "drop")

# Create a bar graph for the total number of deaths
ggplot(total_deaths, aes(x = factor(Year_of_death), y = Total_Deaths, fill = Age_group)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Year of Death",
       y = "Total Number of Deaths") +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  theme_minimal()

```

```{r}
#| label: figure2
#| fig-cap: Deaths Per Year per Gender
#| echo: false
#| warning: false
#| message: false
# Aggregate data to get the total number of deaths for each year and gender
total_deaths_gender <- cleaned_data %>%
  group_by(Year_of_death, Gender) %>%
  summarise(Total_Deaths = sum(Count), .groups = "drop")

# Create a grouped bar graph for male and female
ggplot(total_deaths_gender, aes(x = factor(Year_of_death), y = Total_Deaths, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year of Death",
       y = "Total Number of Deaths") +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62")) +
  theme_minimal()

```
# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

second_model <-
  readRDS(file = here::here("models/second_model.rds"))

third_model <-
  readRDS(file = here::here("models/third_model.rds"))

```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "Gender Only" = second_model,
    "Age Only" = first_model,
    "Age and Gender" = third_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior we implement a posterior predictive check. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay is a trace and rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References